Estimating the socio-environmental impacts of car substitution by bicycle and public transit using open tools

### 2023-09-21 paper/PaperTRA/PaperTRA.Rmd
references paper
move to full paper TRA

---
# Supported options: 
#   sn-nature:       Style for submissions to Nature Portfolio journals
#   sn-basic:        Basic Springer Nature Reference Style/Chemistry Reference Style
#   sn-mathphys:     Math and Physical Sciences Reference Style
#   sn-aps:          American Physical Society (APS) Reference Style
#   sn-vancouver:    Vancouver Reference Style
#   sn-apa:          APA Reference Style 
#   sn-chicago:      Chicago-based Humanities Reference Style
#   default:         Default

classoptions: 
  - sn-apa      
  # - Numbered      # Optional: Use numbered references instead of namedate references (only for sn-nature, sn-basic, sn-vancouver, sn-chicago, sn-mathphys or sn-nature)
  - referee       # Optional: Use double line spacing
  # - lineno        # Optional: Add line numbers
  # - iicol         # Optional: Double column layour

title: Estimating the socio-environmental impacts of car substitution by bicycle and public transit using open tools
# titlerunning: Article Title runing

authors: 
  - firstname: Rosa
    lastname: Félix
    email: rosamfelix@tecnico.ulisboa.pt
    affiliation: 1
    corresponding: TRUE
    equalcont: These authors contributed equally to this work.

  - firstname: Filipe
    lastname: Moura
    email: fmoura@tecnico.ulisboa.pt
    affiliation: 1
    corresponding: FALSE
    equalcont: These authors contributed equally to this work.

  - firstname: Robin
    lastname: Lovelace
    email: r.lovelace@leeds.ac.uk
    affiliation: 2
    equalcont: These authors contributed equally to this work.

affiliations:
  - number: 1
    corresponding: TRUE
    info:
      orgdiv: CERIS - Instituto Superior Técnico
      orgname: University of Lisbon
    address:
        street: Av Rovisco Pais 1
        city: Lisbon
        postcode: 1049-001
        # state: State
        country: Portugal
  - number: 2
    corresponding: FALSE
    info:
      orgdiv: Institute for Transport Studies
      orgname: University of Leeds
    address:
        street: 34-40 University Rd
        city: Leeds
        postcode: LS2 9JT
        # state: State
        country: United Kingdom

keywords:
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods

# Sample for structured abstract - MAX 200 words - ainda tem 274
abstract: |
  In metropolitan areas, car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This approach requires interventions to make cycling more appealing, and the resulting public investments can have significant social and environmental benefits. This paper focuses on estimating the potential for cycling + PT as a substitute for car trips in the Lisbon metropolitan area and assessing its socio-environmental impacts using open data and open source tools.

  To achieve the cycling targets set by the Portuguese cycling strategy, the Department of Transport introduced _biclaR_, a decision support tool that facilitates the design and development of a metropolitan cycling network. A scenario of intermodality introduced, and its socio-environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. Additionally, we estimated the impacts of shifting car trips to PT and monetize them with the EU Guide to cost-benefit analysis.

  The results indicate that 20% of the current trips can be made with the bicycle + PT combination, with an additional 12% of PT trips being potentially replaced. Shifting to cycling for the first-and-last mile can reduce annual CO2eq emissions by 6,000 to 15,000 tons/day, and the 10-year socio-environmental benefits account for €230 to €590 million, depending on the cycling targets. For the PT leg, the transfer from car results in the avoidance of 8,500 to 20,800 tons of CO2eq emissions per year.

  The provided information on socio-economic benefits can support policymakers in prioritizing interventions to reduce the reliance on individual motorized transportation and effectively communicate their decisions.

bibliography: bibliography.bib

header-includes: |

output: rticles::springer_article
---

@@

# Introduction {#sec1}

The Introduction section should explain the background to the study, its aims, a summary of the existing literature and why this study was necessary.

@@

The intermodality scenario considers trips that can combine PT and cycling for the first-and-last legs. Conservatively, we considered the sum of first-and-last legs up to 5 km. Furthermore, we restricted PT use to unimodal trips without transfers (although they can be included in future modeling). Finally, we only included PT modes that can practically accommodate bicycles, such as trains, ferries, trams, and intermunicipal bus lines with bike racks (Figure 1).

@@

@@

@@

@@

<!-- The Introduction section, of referenced text @bib1 expands on the background of the work (some overlap with the Abstract is acceptable). The introduction should not include subheadings. -->

# Methods {#sec2}

The methods section should include:

* the aim, design and setting of the study
* the characteristics of participants or description of materials
* a clear description of all processes and methodologies employed. Generic names should generally be used. When proprietary * brands are used in research, include the brand names in parentheses
* the type of statistical analysis used, including a power calculation if appropriate
* studies involving human participants, data or tissue or animals must include statement on ethics approval and consent

```{r fig1, fig.cap = "This is an example of a caption", fig.width = 5}
# plot(pressure)

```

# Results and Discussion {#sec11}

Topical subheadings are allowed. Authors must ensure that their Methods section includes adequate experimental and characterization data necessary for others in the field to reproduce their work. Authors are encouraged to include RIIDs where appropriate.

**Ethical approval declarations** (only required where applicable) Any article reporting experiment/s carried out on (i)~live vertebrate (or higher invertebrates), (ii)~humans or (iii)~human samples must include an unambiguous statement within the methods section that meets the following requirements:

1. Approval: a statement which confirms that all experimental protocols were approved by a named institutional and/or licensing committee. Please identify the approving body in the methods section

2. Accordance: a statement explicitly saying that the methods were carried out in accordance with the relevant guidelines and regulations

3. Informed consent (for experiments involving humans or human tissue samples): include a statement confirming that informed consent was obtained from all participants and/or their legal guardian/s

If your manuscript includes potentially identifying patient/participant information, or if it describes human transplantation research, or if it reports results of a clinical trial then  additional information will be required. Please visit (<https://www.nature.com/nature-research/editorial-policies>) for Nature Portfolio journals, (<https://www.springer.com/gp/authors-editors/journal-author/journal-author-helpdesk/publishing-ethics/14214>) for Springer Nature journals, or (<https://www.biomedcentral.com/getpublished/editorial-policies\#ethics+and+consent>) for BMC.

# Conclusion {#sec13}

This should state clearly the main conclusions and provide an explanation of the importance and relevance of the study to the field.

Conclusions may be used to restate your hypothesis or research question, restate your major findings, explain the relevance and the added value of your work, highlight any limitations of your study, describe future directions for research and recommendations.

In some disciplines use of Discussion or 'Conclusion' is interchangeable. It is not mandatory to use both. Please refer to Journal-level guidance for any specific requirements.

\backmatter

\bmhead{Acknowledgments}

Acknowledgments are not compulsory. Where included they should be brief. Grant or contribution numbers may be acknowledged.

Please refer to Journal-level guidance for any specific requirements.

# Declarations {.unnumbered}

Some journals require declarations to be submitted in a standardised format. Please check the Instructions for Authors of the journal to which you are submitting to see if you need to complete this section. If yes, your manuscript must contain the following sections under the heading `Declarations':

* Funding
* Conflict of interest/Competing interests (check journal-specific guidelines for which heading to use)
* Ethics approval
* Consent to participate
* Consent for publication
* Availability of data and materials
* Code availability
* Authors' contributions

\noindent
If any of the sections are not relevant to your manuscript, please include the heading and write `Not applicable' for that section.

>>>>

---
classoption: "runningheads"
title: "Estimating the socio-environmental impacts of car substitution by bicycle and public transit using open tools"
shorttitle: "Estimating the SE impacts of car substitution by bicycle and public transit"

authors: 
  - name: "Rosa Félix"
    inst: "1"
    orcid: "0000-0002-5642-6006"
  - name: "Filipe Moura"
    inst: "1"
    orcid: "0000-0001-7749-8490"
  - name: "Robin Lovelace"
    inst: "2"
    orcid: "0000-0001-5679-6536"
# First names are abbreviated in the running head.
# If there are more than two authors, 'et al.' is used.
authorrunning: "R. Félix et al." 
institutes: 
  - name: "CERIS - Instituto Superior Técnico, University of Lisbon. Av Rovisco Pais 1049-001 Lisboa, Portugal"
    email: "rosamfelix@tecnico.ulisboa.pt"
  - name: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"

abstract: | 
  In metropolitan areas, car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper focuses on estimating the potential for cycling + PT as a substitute for car trips in the Lisbon metropolitan area and assessing its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). A scenario of intermodality introduced, and its socio-environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. Additionally, the impacts of shifting car trips to PT were estimated and monetized.
  The results indicate that 20% of the current trips can be made with the bicycle + PT combination. Shifting to cycling for the first-and-last mile can reduce annual CO2eq emissions from 6,000 tons/day, and the 10-year socio-environmental benefits account from €230 million. For the PT leg, the transfer from car results in the avoidance of at lest 8,500 tons of CO2eq emissions per year.
  The information on socio-economic benefits can support policymakers in prioritizing interventions to reduce the reliance on individual motorized transportation and effectively communicate their decisions.
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
bibliography: bibliography.bib
# Use csl when using default citation_package for Pandoc citeproc
# csl: https://www.zotero.org/styles/springer-lecture-notes-in-computer-science
# citations with arabic numbers is prefered when using natbib
natbiboptions: numbers

header-includes:
  - \usepackage[hidelinks]{hyperref}

output: 
  rticles::lncs_article:
    citation_package: default # or natbib
---

@@

# Introduction

**full paper**: 4-6 pages in length (typically up to 3,000 words).

@@

This research aims to present and discuss the methods used to estimate…

Propensity to Cycle Tool

adding up an intermodality scenario to estimate cycling potential to public transit interfaces, and thus to support planning and prioritize investments in the cycling network.

# Methods

## Case Study

## Modeling Origin-Destination trips

## Modeling intermodality

The intermodality scenario considers trips that can combine PT and cycling for the first-and-last legs. Conservatively, we considered the sum of first-and-last legs up to 5 km. Furthermore, we restricted PT use to unimodal trips without transfers (although they can be included in future modeling). [@bib2, @felix2023]

Finally, we only included PT modes that can practically accommodate bicycles, such as trains, ferries, trams, and intermunicipal bus lines with bike racks (\ \ref{fig:map1}).

```{r map1, fig.cap="Interfaces and lines considered, by transport mode, in the Lisbon metropolitan area"}
plot(pressure)
```

@@

## Assessing socio-environmental benefits

@@

Table\ \ref{tab:table} gives a summary of all heading levels.

```{r table}
knitr::kable(head(pressure), 
             format = "latex",
             caption = "Table captions should be placed above the tables.")
```

# Results and Discussion

@@

Map result (\ \ref{fig:map2}).

```{r map2, fig.cap="Bike routes with highest potential to serve as first and last mile when replacing cycling and PT from car trips (screenshot of the interactive online tool)."}
plot(pressure)
```

# Conclusion

@@

### Acknowledgements {.unnumbered}
Please place your acknowledgments at the end of the paper, preceded by an unnumbered run-in heading (i.e. 3rd-level heading).

Thomas Götshi - HAAS.

### 2023-09-21 paper/PaperTRA/PaperTRA.tex
move to full paper TRA

%Version 2.1 April 2023
% See section 11 of the User Manual for version history
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[sn-apa,referee,pdflatex]{sn-jnl}

%%%% Standard Packages
%%<additional latex packages if required can be included here>

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%
%%%%

%%%%%=============================================================================%%%%
%%%%  Remarks: This template is provided to aid authors with the preparation
%%%%  of original research articles intended for submission to journals published
%%%%  by Springer Nature. The guidance has been prepared in partnership with
%%%%  production teams to conform to Springer Nature technical requirements.
%%%%  Editorial and presentation requirements differ among journal portfolios and
%%%%  research disciplines. You may find sections in this template are irrelevant
%%%%  to your work and are empowered to omit any such section if allowed by the
%%%%  journal you intend to submit to. The submission guidelines and policies
%%%%  of the journal take precedence. A detailed User Manual is available in the
%%%%  template package for technical guidance.
%%%%%=============================================================================%%%%

\raggedbottom

@@

@@

\title[]{Estimating the socio-environmental impacts of car substitution
by bicycle and public transit using open tools}

%%=============================================================%%
%% Prefix	-> \pfx{Dr}
%% GivenName	-> \fnm{Joergen W.}
%% Particle	-> \spfx{van der} -> surname prefix
%% FamilyName	-> \sur{Ploeg}
%% Suffix	-> \sfx{IV}
%% NatureName	-> \tanm{Poet Laureate} -> Title after name
%% Degrees	-> \dgr{MSc, PhD}
%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate}
%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
%%=============================================================%%

\author*[1]{\fnm{Rosa} \sur{Félix} }\email{\href{mailto:rosamfelix@tecnico.ulisboa.pt}{\nolinkurl{rosamfelix@tecnico.ulisboa.pt}}}

\author[1]{\fnm{Filipe} \sur{Moura} }\email{\href{mailto:fmoura@tecnico.ulisboa.pt}{\nolinkurl{fmoura@tecnico.ulisboa.pt}}}

\author[2]{\fnm{Robin} \sur{Lovelace} }\email{\href{mailto:r.lovelace@leeds.ac.uk}{\nolinkurl{r.lovelace@leeds.ac.uk}}}

  \affil*[1]{\orgdiv{CERIS - Instituto Superior
Técnico}, \orgname{University of
Lisbon}, \orgaddress{\city{Lisbon}, \country{Portugal}, \postcode{1049-001}, \street{Av
Rovisco Pais 1}}}
  \affil[2]{\orgdiv{Institute for Transport
Studies}, \orgname{University of
Leeds}, \orgaddress{\city{Leeds}, \country{United
Kingdom}, \postcode{LS2 9JT}, \street{34-40 University Rd}}}

\abstract{In metropolitan areas, car trips can be replaced by a
combination of public transit and cycling for the first-and-last mile.
This approach requires interventions to make cycling more appealing, and
the resulting public investments can have significant social and
environmental benefits. This paper focuses on estimating the potential
for cycling + PT as a substitute for car trips in the Lisbon
metropolitan area and assessing its socio-environmental impacts using
open data and open source tools.

To achieve the cycling targets set by the Portuguese cycling strategy,
the Department of Transport introduced \emph{biclaR}, a decision support
tool that facilitates the design and development of a metropolitan
cycling network. A scenario of intermodality introduced, and its
socio-environmental impacts were assessed using the \emph{HEAT for
Cycling} and the \emph{HEAT as a Service} tools. Additionally, we
estimated the impacts of shifting car trips to PT and monetize them with
the EU Guide to cost-benefit analysis.

The results indicate that 20\% of the current trips can be made with the
bicycle + PT combination, with an additional 12\% of PT trips being
potentially replaced. Shifting to cycling for the first-and-last mile
can reduce annual CO2eq emissions by 6,000 to 15,000 tons/day, and the
10-year socio-environmental benefits account for €230 to €590 million,
depending on the cycling targets. For the PT leg, the transfer from car
results in the avoidance of 8,500 to 20,800 tons of CO2eq emissions per
year.

The provided information on socio-economic benefits can support
policymakers in prioritizing interventions to reduce the reliance on
individual motorized transportation and effectively communicate their
decisions.}

\keywords{Active transport, Intermodality, First and last mile, Health
economic assessment, Environmental impacts, Open data and methods}

\maketitle

\hypertarget{sec1}{%
\section{Introduction}\label{sec1}}

The Introduction section should explain the background to the study, its
aims, a summary of the existing literature and why this study was
necessary.

@@

The intermodality scenario considers trips that can combine PT and
cycling for the first-and-last legs. Conservatively, we considered the
sum of first-and-last legs up to 5 km. Furthermore, we restricted PT use
to unimodal trips without transfers (although they can be included in
future modeling). Finally, we only included PT modes that can
practically accommodate bicycles, such as trains, ferries, trams, and
intermunicipal bus lines with bike racks (Figure 1).

@@

@@

@@

@@

The Introduction section, of referenced text \citet{bib1} expands on the
background of the work (some overlap with the Abstract is acceptable).
The introduction should not include subheadings.

Springer Nature does not impose a strict layout as standard however
authors are advised to check the individual requirements for the journal
they are planning to submit to as there may be journal-level
preferences. When preparing your text please also be aware that some
stylistic choices are not supported in full text XML (publication
version), including coloured font. These will not be replicated in the
typeset article if it is accepted.

\hypertarget{sec2}{%
\section{Methods}\label{sec2}}

The methods section should include:

\begin{itemize}
\tightlist
\item
  the aim, design and setting of the study
\item
  the characteristics of participants or description of materials
\item
  a clear description of all processes and methodologies employed.
  Generic names should generally be used. When proprietary * brands are
  used in research, include the brand names in parentheses
\item
  the type of statistical analysis used, including a power calculation
  if appropriate
\item
  studies involving human participants, data or tissue or animals must
  include statement on ethics approval and consent
\end{itemize}

\begin{figure}
\includegraphics{PaperTRA_files/figure-latex/fig1-1} \caption{This is an example of a caption}\label{fig:fig1}
\end{figure}

\hypertarget{sec11}{%
\section{Results and Discussion}\label{sec11}}

Topical subheadings are allowed. Authors must ensure that their Methods
section includes adequate experimental and characterization data
necessary for others in the field to reproduce their work. Authors are
encouraged to include RIIDs where appropriate.

\textbf{Ethical approval declarations} (only required where applicable)
Any article reporting experiment/s carried out on
(i)\textasciitilde live vertebrate (or higher invertebrates),
(ii)\textasciitilde humans or (iii)\textasciitilde human samples must
include an unambiguous statement within the methods section that meets
the following requirements:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Approval: a statement which confirms that all experimental protocols
  were approved by a named institutional and/or licensing committee.
  Please identify the approving body in the methods section
\item
  Accordance: a statement explicitly saying that the methods were
  carried out in accordance with the relevant guidelines and regulations
\item
  Informed consent (for experiments involving humans or human tissue
  samples): include a statement confirming that informed consent was
  obtained from all participants and/or their legal guardian/s
\end{enumerate}

If your manuscript includes potentially identifying patient/participant
information, or if it describes human transplantation research, or if it
reports results of a clinical trial then additional information will be
required. Please visit
(\url{https://www.nature.com/nature-research/editorial-policies}) for
Nature Portfolio journals,
(\url{https://www.springer.com/gp/authors-editors/journal-author/journal-author-helpdesk/publishing-ethics/14214})
for Springer Nature journals, or
(\url{https://www.biomedcentral.com/getpublished/editorial-policies/\#ethics+and+consent})
for BMC.

\hypertarget{sec13}{%
\section{Conclusion}\label{sec13}}

This should state clearly the main conclusions and provide an
explanation of the importance and relevance of the study to the field.

Conclusions may be used to restate your hypothesis or research question,
restate your major findings, explain the relevance and the added value
of your work, highlight any limitations of your study, describe future
directions for research and recommendations.

In some disciplines use of Discussion or `Conclusion' is
interchangeable. It is not mandatory to use both. Please refer to
Journal-level guidance for any specific requirements.

\backmatter

\bmhead{Acknowledgments}

Acknowledgments are not compulsory. Where included they should be brief.
Grant or contribution numbers may be acknowledged.

Please refer to Journal-level guidance for any specific requirements.

\hypertarget{declarations}{%
\section*{Declarations}\label{declarations}}
\addcontentsline{toc}{section}{Declarations}

Some journals require declarations to be submitted in a standardised
format. Please check the Instructions for Authors of the journal to
which you are submitting to see if you need to complete this section. If
yes, your manuscript must contain the following sections under the
heading `Declarations':

\begin{itemize}
\tightlist
\item
  Funding
\item
  Conflict of interest/Competing interests (check journal-specific
  guidelines for which heading to use)
\item
  Ethics approval
\item
  Consent to participate
\item
  Consent for publication
\item
  Availability of data and materials
\item
  Code availability
\item
  Authors' contributions
\end{itemize}

\noindent If any of the sections are not relevant to your manuscript,
please include the heading and write `Not applicable' for that section.

\renewcommand\refname{References}
\bibliography{bibliography.bib}

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.

@@

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[hidelinks]{hyperref}

\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage{hyperref}
\usepackage{color}
\renewcommand\UrlFont{\color{blue}\rmfamily}

@@

\title{Estimating the socio-environmental impacts of car substitution by
bicycle and public transit using open tools}
%
\titlerunning{Estimating the SE impacts of car substitution by bicycle
and public transit}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Rosa Félix\inst{1}\orcidID{0000-0002-5642-6006} \and Filipe
Moura\inst{1}\orcidID{0000-0001-7749-8490} \and Robin
Lovelace\inst{2}\orcidID{0000-0001-5679-6536}}

\authorrunning{R. Félix et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%

\institute{CERIS - Instituto Superior Técnico, University of Lisbon. Av
Rovisco Pais 1049-001 Lisboa, Portugal\\
\email{\href{mailto:rosamfelix@tecnico.ulisboa.pt}{\nolinkurl{rosamfelix@tecnico.ulisboa.pt}}}\\ \and Institute
for Transport Studies, University of Leeds. 34-40 University Rd, Leeds
LS2 9JT, UK}

\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In metropolitan areas, car trips can be replaced by a combination of
public transit and cycling for the first-and-last mile. This paper
focuses on estimating the potential for cycling + PT as a substitute for
car trips in the Lisbon metropolitan area and assessing its
socio-environmental impacts using open data and open source tools. A
decision support tool that facilitates the design and development of a
metropolitan cycling network was developed (\emph{biclaR}). A scenario
of intermodality introduced, and its socio-environmental impacts were
assessed using the \emph{HEAT for Cycling} and the \emph{HEAT as a
Service} tools. Additionally, the impacts of shifting car trips to PT
were estimated and monetized. The results indicate that 20\% of the
current trips can be made with the bicycle + PT combination. Shifting to
cycling for the first-and-last mile can reduce annual CO2eq emissions
from 6,000 tons/day, and the 10-year socio-environmental benefits
account from €230 million. For the PT leg, the transfer from car results
in the avoidance of at lest 8,500 tons of CO2eq emissions per year. The
information on socio-economic benefits can support policymakers in
prioritizing interventions to reduce the reliance on individual
motorized transportation and effectively communicate their decisions.

\keywords{Active transport \and Intermodality \and First and last
mile \and Health economic assessment \and Environmental
impacts \and Open data and methods}

\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

This research aims to present and discuss the methods used to
estimate\ldots{}

Propensity to Cycle Tool

adding up an intermodality scenario to estimate cycling potential to
public transit interfaces, and thus to support planning and prioritize
investments in the cycling network.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{case-study}{%
\subsection{Case Study}\label{case-study}}

\hypertarget{modeling-origin-destination-trips}{%
\subsection{Modeling Origin-Destination
trips}\label{modeling-origin-destination-trips}}

\hypertarget{modeling-intermodality}{%
\subsection{Modeling intermodality}\label{modeling-intermodality}}

The intermodality scenario considers trips that can combine PT and
cycling for the first-and-last legs. Conservatively, we considered the
sum of first-and-last legs up to 5 km. Furthermore, we restricted PT use
to unimodal trips without transfers (although they can be included in
future modeling). Félix, Lovelace, and Moura (2022)

Finally, we only included PT modes that can practically accommodate
bicycles, such as trains, ferries, trams, and intermunicipal bus lines
with bike racks (~\ref{fig:map1}).

\begin{figure}
\includegraphics{PaperTRA_files/figure-latex/map1-1} \caption{Interfaces and lines considered, by transport mode, in the Lisbon metropolitan area}\label{fig:map1}
\end{figure}

@@

\hypertarget{assessing-socio-environmental-benefits}{%
\subsection{Assessing socio-environmental
benefits}\label{assessing-socio-environmental-benefits}}

@@

Table~\ref{tab:table} gives a summary of all heading levels.

\begin{table}

\caption{\label{tab:table}Table captions should be placed above the tables.}
\centering
\begin{tabular}[t]{r|r}
\hline
temperature & pressure\\
\hline
0 & 0.0002\\
\hline
20 & 0.0012\\
\hline
40 & 0.0060\\
\hline
60 & 0.0300\\
\hline
80 & 0.0900\\
\hline
100 & 0.2700\\
\hline
\end{tabular}
\end{table}

\hypertarget{results-and-discussion}{%
\section{Results and Discussion}\label{results-and-discussion}}

@@

Map result (~\ref{fig:map2}).

\begin{figure}
\includegraphics{PaperTRA_files/figure-latex/map2-1} \caption{Bike routes with highest potential to serve as first and last mile when replacing cycling and PT from car trips (screenshot of the interactive online tool).}\label{fig:map2}
\end{figure}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

@@

\hypertarget{acknowledgements}{%
\subsubsection*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{subsubsection}{Acknowledgements}

Please place your acknowledgments at the end of the paper, preceded by
an unnumbered run-in heading (i.e.~3rd-level heading).

Thomas Götshi - HAAS.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-felix2023}{}}%
Félix, Rosa, Robin Lovelace, and Filipe Moura. 2022. {``{biclaR -
Ferramenta de apoio ao planeamento da rede ciclável na área
metropolitana de Lisboa}.''} {CERIS - Instituto Superior Técnico and
Transportes Metropolitanos de Lisboa}.
\url{https://biclar.tmlmobilidade.pt}.

\leavevmode\vadjust pre{\hypertarget{ref-bib2}{}}%
Slifka, M. K., and J. L. Whitton. 2000. {``Clinical Implications of
Dysregulated Cytokine Production.''} \emph{J. {M}ol. {M}ed.} 78: 74--80.
\url{https://doi.org/10.1007/s001090000086}.

\end{CSLReferences}

%
% ---- Bibliography ----
